Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-09-15 prop_letter.tex
prop_letter.tex \section{Gleason} Updating...

The quantum probability measure~$\mu_{1}$ maps some projectors to
$0$ or $0.5$, which are replaced by $\left[0,0\right]$ or $\left[0.5,0.5\right]$
to get $\bar{\mu}_{3}$. This is replacing two numbers by the smallest
intervals containing them. In general, any set of the numbers can
be replaced by an interval containing them if we carefully pick a
set of intervals and the rule to select the correct interval when
intervals are overlapping. For example, consider the following collection
of intervals: 
\begin{eqnarray}
\mathscr{I}_{n} & = & \set{\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\nonumber \\
 &  & \cup\{\imposs,\necess\}
\end{eqnarray}
which are overlapping intervals with length~$1/n$ except the intervals
at length~$\left(2n\right)^{-1}$ at each end. Since these intervals
are overlapping, we define the interior part of these intervals

Then, we can define a QIVPM mapping to $\mathscr{I}_{n}$ based on
any Born rule probability~$\mu_{\rho}^{\mathrm{B}}$ on a three-dimensional
Hilbert space by replacing the probability~$\mu_{\rho}^{\mathrm{B}}\left(P\right)$
with the interval containing it, $\iota_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)$.
Since intervals in $\mathscr{I}_{n}$ are overlapping, if $\mu_{\rho}^{\mathrm{B}}\left(P\right)$
are contained in both intervals~$[\ell_{1},r_{1}]$ and $[\ell_{2},r_{2}]$,
$\iota_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)$ will
be $[\ell_{i},r_{i}]$ if $\iota_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)$
is closer to the center of $[\ell_{i},r_{i}]$\yutsung{Verify...}.
When there is a tie, we set $\iota_{n}\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)=\left[\frac{k-1}{2n},\frac{k+1}{2n}\right]\cap\left[0,1\right]$
for even $k$. For example, when $n=2$, 
\begin{eqnarray}
\mathscr{I}_{2} & = & \set{\left[\tfrac{k-1}{4},\tfrac{k+1}{4}\right]\cap\left[0,1\right]}{k=0,\ldots,4}\nonumber \\
 &  & \cup\{\imposs,\necess\}\nonumber \\
 & = & \left\{ \imposs,\left[0,\tfrac{1}{4}\right],\left[0,{\textstyle \tfrac{1}{2}}\right],\left[\tfrac{1}{4},\tfrac{3}{4}\right],\right.\nonumber \\
 &  & \left.\left[\tfrac{1}{2},1\right],\left[\tfrac{1}{4},1\right],\necess\right\} \,.
\end{eqnarray}
It is straightforward to verify $\iota_{2}\circ\mu_{1}$ and $\iota_{2}\circ\mu_{\mathbb{1}/3}^{\mathrm{B}}$
are QIVPMs whose values on the minimal subspace of events containing
$\proj{0}$, $\proj{1}$, and $\proj{2}$, $\events_{C}$, is computed
by Table~\ref{tab:Compute-the-QIVPM}. 
\begin{table}
\caption{\label{tab:Compute-the-QIVPM}Compute $\iota_{2}\circ\mu_{1}$ and
$\iota_{2}\circ\mu_{\mathbb{1}/3}^{\mathrm{B}}$ based on $\mu_{1}$
and $\mu_{\mathbb{1}/3}^{\mathrm{B}}$ on $\events_{C}$.}
\begin{tabular}{ccccc}
\toprule 
$P$ & $\mu_{1}\left(P\right)$ & $\iota_{2}\left(\mu_{1}\left(P\right)\right)$ & $\mu_{\mathbb{1}/3}^{\mathrm{B}}\left(P\right)$ & $\iota_{2}\left(\mu_{\mathbb{1}/3}^{\mathrm{B}}\left(P\right)\right)$\tabularnewline
\midrule
\midrule 
$\mathbb{0}$ & $0$ & $\left[0,0\right]$ & $0$ & $\left[0,0\right]$\tabularnewline
\midrule 
$\proj{0}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$ & $\frac{1}{3}$ & $\left[0,\frac{1}{2}\right]$\tabularnewline
\midrule 
$\proj{1}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$ & $\frac{1}{3}$ & $\left[0,\frac{1}{2}\right]$\tabularnewline
\midrule 
$\proj{2}$ & $0$ & $\left[0,0\right]$ & $\frac{1}{3}$ & $\left[0,\frac{1}{2}\right]$\tabularnewline
\midrule 
$\mathbb{1}-\proj{0}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$ & $\frac{2}{3}$ & $\left[\frac{1}{2},1\right]$\tabularnewline
\midrule 
$\mathbb{1}-\proj{1}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$ & $\frac{2}{3}$ & $\left[\frac{1}{2},1\right]$\tabularnewline
\midrule 
$\mathbb{1}-\proj{2}$ & $1$ & $\left[1,1\right]$ & $\frac{2}{3}$ & $\left[\frac{1}{2},1\right]$\tabularnewline
\midrule 
$\mathbb{1}$ & $1$ & $\left[1,1\right]$ & $1$ & $\left[1,1\right]$\tabularnewline
\bottomrule
\end{tabular}
\end{table}

Given any state~$\rho$ satisfying $0<\mu_{\rho}^{\mathrm{B}}\left(P\right)<1$
for any $P\in\events\backslash\left\{ \mathbb{0},\mathbb{1}\right\} $,
called totally mixed state, we know not only $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
is a QIVPM, but also the collection of states consistent it get closer
and closer to $\rho$ as the intervals get narrower and narrower,
eventually collapsing to $\rho$ when the intervals become infinitely
precise real numbers. To state this result, we need to define a notion
of distance based on the following norm of a matrix~$A$ as $\left\Vert A\right\Vert =\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{A}{\psi}\right|$~\citep{544199}.
If a state~$\rho'$ is consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$,
its distance with $\rho$, $\left\Vert \rho-\rho'\right\Vert $, must
be bounded by $1/\left(12n\right)$. Reversely, if the distance between
$\rho$ and $\rho'$ is bounded by $1/n$, then $\rho'$ is consistent
with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$. In other words, for
the states consistent with this QIVPM, $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
will become closer and closer to $\rho'$ as $n$ goes to infinity.

For example, since $\mu_{1}\left(\proj{2}\right)=0$, any state corresponds
to $\mu_{1}$is not totally mixed state. On the other hand, $\mathbb{1}/3$
is a totally mixed state. \yutsung{{To be continue...}}

>>>>

The quantum probability measure~$\mu_{1}$ maps some projectors to
$0$ or $0.5$, which are replaced by $\left[0,0\right]$ or $\left[0.5,0.5\right]$
to get $\bar{\mu}_{3}$. The way we replace two numbers by the smallest
intervals containing them can be generalized if we carefully pick
a set of intervals and the rule to select the correct interval when
intervals are overlapping. For example, consider the following collection
of intervals: 
\begin{eqnarray}
\mathscr{I}_{n} & = & \set{\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\nonumber \\
 &  & \cup\{\imposs,\necess\}
\end{eqnarray}
which are overlapping intervals with length~$1/n$ except the intervals
at length~$\left(2n\right)^{-1}$ at each end. Since these intervals
are overlapping, we define the non-overlapping interior part of a
interval according to the following rule.\begin{subequations}
\begin{eqnarray}
\textrm{Int}\left(\imposs\right) & = & \imposs\,,\\
\textrm{Int}\left(\left[0,\tfrac{1}{2n}\right]\right) & = & \left(0,\tfrac{0.5}{2n}\right]\,,\\
\textrm{Int}\left(\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]\right) & = & \left[\tfrac{k-0.5}{2n},\tfrac{k+0.5}{2n}\right]\nonumber \\
 &  & \textrm{for even }0<k<2n\,,\\
\textrm{Int}\left(\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]\right) & = & \left(\tfrac{k-0.5}{2n},\tfrac{k+0.5}{2n}\right)\nonumber \\
 &  & \textrm{for odd }k\,,\\
\textrm{Int}\left(\left[\tfrac{2n-1}{2n},1\right]\right) & = & \left[\tfrac{2n-0.5}{2n},1\right)\,,\\
\textrm{Int}\left(\necess\right) & = & \necess\,.
\end{eqnarray}
\end{subequations}Hence, given a real-valued probability~$x$, we
will replace it with an interval~$\iota_{n}\left(x\right)=[\ell,r]$
if $x\in\textrm{Int}\left([\ell,r]\right)$.

After we have a rule to replace a real number by an interval, any
Born rule probability~$\mu_{\rho}^{\mathrm{B}}$ on a three-dimensional
Hilbert space can be extended to a QIVPM~$\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$,
where $\circ$ is the function concatenation. For example, when $n=2$,
we can summarize the set of intervals~$\mathscr{I}_{2}$ and their
interior in Table~\ref{tab:I2}.
\begin{table}
\caption{\label{tab:I2}Intervals in $\mathscr{I}_{2}$ and their interior.}
\begin{tabular}{cc}
\toprule 
$[\ell,r]$ & $\textrm{Int}\left([\ell,r]\right)$\tabularnewline
\midrule
\midrule 
$\imposs$ & $\imposs$\tabularnewline
\midrule 
$\left[0,\tfrac{1}{4}\right]$ & $\left(0,\tfrac{1}{8}\right]$\tabularnewline
\midrule 
$\left[0,{\textstyle \tfrac{1}{2}}\right]$ & $\left(\tfrac{1}{8},\tfrac{3}{8}\right)$\tabularnewline
\midrule 
$\left[\tfrac{1}{4},\tfrac{3}{4}\right]$ & $\left[\tfrac{3}{8},\tfrac{5}{8}\right]$\tabularnewline
\midrule 
$\left[\tfrac{1}{2},1\right]$ & $\left(\tfrac{5}{8},\tfrac{7}{8}\right)$\tabularnewline
\midrule 
$\left[\tfrac{1}{4},1\right]$ & $\left[\tfrac{7}{8},1\right)$\tabularnewline
\midrule 
$\necess$ & $\necess$\tabularnewline
\bottomrule
\end{tabular}
\end{table}
Given two quantum probability measure~$\mu_{1}$ and $\mu_{\mathbb{1}/3}^{\mathrm{B}}$,
where $\frac{\mathbb{1}}{3}=\frac{1}{3}\left(\proj{0}+\proj{1}+\proj{2}\right)$,
it is then easy to compute $\iota_{2}\circ\mu_{1}$ and $\iota_{2}\circ\mu_{\mathbb{1}/3}^{\mathrm{B}}$
on the minimal subspace of events containing $\proj{0}$, $\proj{1}$,
and $\proj{2}$, $\events_{C}$ as in Table~\ref{tab:Compute-the-QIVPM}.
\begin{table}
\caption{\label{tab:Compute-the-QIVPM}Compute $\iota_{2}\circ\mu_{1}$ and
$\iota_{2}\circ\mu_{\mathbb{1}/3}^{\mathrm{B}}$ based on $\mu_{1}$
and $\mu_{\mathbb{1}/3}^{\mathrm{B}}$ on $\events_{C}$.}
\begin{tabular}{ccccc}
\toprule 
$P$ & $\mu_{1}\left(P\right)$ & $\iota_{2}\left(\mu_{1}\left(P\right)\right)$ & $\mu_{\mathbb{1}/3}^{\mathrm{B}}\left(P\right)$ & $\iota_{2}\left(\mu_{\mathbb{1}/3}^{\mathrm{B}}\left(P\right)\right)$\tabularnewline
\midrule
\midrule 
$\mathbb{0}$ & $0$ & $\imposs$ & $0$ & $\imposs$\tabularnewline
\midrule 
$\proj{0}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$ & $\frac{1}{3}$ & $\left[0,\frac{1}{2}\right]$\tabularnewline
\midrule 
$\proj{1}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$ & $\frac{1}{3}$ & $\left[0,\frac{1}{2}\right]$\tabularnewline
\midrule 
$\proj{2}$ & $0$ & $\imposs$ & $\frac{1}{3}$ & $\left[0,\frac{1}{2}\right]$\tabularnewline
\midrule 
$\mathbb{1}-\proj{0}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$ & $\frac{2}{3}$ & $\left[\frac{1}{2},1\right]$\tabularnewline
\midrule 
$\mathbb{1}-\proj{1}$ & $0.5$ & $\left[\frac{1}{4},\frac{3}{4}\right]$ & $\frac{2}{3}$ & $\left[\frac{1}{2},1\right]$\tabularnewline
\midrule 
$\mathbb{1}-\proj{2}$ & $1$ & $\necess$ & $\frac{2}{3}$ & $\left[\frac{1}{2},1\right]$\tabularnewline
\midrule 
$\mathbb{1}$ & $1$ & $\necess$ & $1$ & $\necess$\tabularnewline
\bottomrule
\end{tabular}
\end{table}
Verifying $\iota_{2}\circ\mu_{1}$ and $\iota_{2}\circ\mu_{\mathbb{1}/3}^{\mathrm{B}}$
are QIVPMs are also straightforward.

Given any state~$\rho$ satisfying $0<\mu_{\rho}^{\mathrm{B}}\left(P\right)<1$
for any $P\in\events\backslash\left\{ \mathbb{0},\mathbb{1}\right\} $,
called totally mixed state, we know not only $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
is a QIVPM, but also the collection of states consistent it get closer
and closer to $\rho$ as the intervals get narrower and narrower,
eventually collapsing to $\rho$ when the intervals become infinitely
precise real numbers. To state this result, we need to define a notion
of distance based on the following norm of a matrix~$A$ as $\left\Vert A\right\Vert =\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{A}{\psi}\right|$~\citep{544199}.
If a state~$\rho'$ is consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$,
its distance with $\rho$, $\left\Vert \rho-\rho'\right\Vert $, must
be bounded by $\left(12n\right)^{-1}$. Reversely, if the distance
between $\rho$ and $\rho'$ is bounded by $1/n$, then $\rho'$ is
consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$. In other
words, for the states consistent with this QIVPM, $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
will become closer and closer to $\rho'$ as $n$ goes to infinity.

For example, since $\mu_{1}\left(\proj{2}\right)=0$, any state corresponds
to $\mu_{1}$ is not totally mixed state. On the other hand, $\mathbb{1}/3$
is a totally mixed state. \yutsung{To be continue...}
